Comparing vector-based and ACT-R memory models using large-scale datasets:\protect\\User-customized hashtag and tag prediction on Twitter and StackOverflow

### 2014-10-15 writeups/dissertation/dissertation.tex
foo

\abstract{
  The growth of social media and user-created content on online sites provides unique opportunities to study models of declarative memory.
  The tasks of choosing a hashtag for a tweet and tagging a post on StackOverflow were framed as declarative memory retrieval problems.
  Two state-of-the-art cognitively-plausible memory models were evaluated on how accurate they predict a user's chosen tags: an ACT-R based Bayesian model and a random permutation vector-based model.
  Millions of posts and tweets were collected, and two declarative memory models were used to predict hashtags and tags that users choose on Twitter and StackOverflow.
  The results show that past user behavior of tag use is a strong predictor of future behavior, and this past behavior was successfully incorporated into the random permutation model that previously used only context.
  Also, ACT-R's attentional weight term was linked to a common entropy-weighting natural language processing method used to attenuate low-predictor words.
  Word order was not found to be strong predictor of tag use, and the random permutation model performed comparable to the Bayesian model without including word order.
  This shows that the strength of the random permutation model is not in the ability to represent word order, but rather in the way in which context information is successfully compressed.
  Finally, model accuracy was moderate to high for the tasks, which supports the theory that choosing tags on StackOverflow and Twitter is primarily a declarative memory retrieval process. 
  The results of the large-scale exploration show how the architecture of the two memory models can be modified to significantly improve accuracy,
  and may suggest task-independent general modifications that can help improve model fit to human data in a much wider range of domains.
}

@@

The ACT-R cognitive architecture \parencite{Anderson2007} contains one of the leading theories of human cognitive processing.
The declarative memory component in ACT-R is a Bayesian statistical model.
This is the same type of co-occurrence Naive Bayes statistical model that is prominent in the more general machine learning field.
When these Bayesian models are used for machine learning, it is common to use a large number of observations to construct the co-occurrence matrix that is the foundation for the model.
The primary reason for this is that the stability of statistical models increase as sample size increases.

However, it is much less common in the cognitive modeling community to use a large number of observations to build up declarative memory when creating ACT-R models.
If a cognitive modeler is using the declarative memory component of ACT-R,
it is common to either construct the co-occurrences by hand, or use a small training set of data to initially build up the declarative memory system.
This approach can and has worked well.
Essentially the modeler is only including the observations in declarative memory that are relevant to the specific task that they are studying.

@@

This has important implications for understanding declarative memory:
First, it is difficult to verify and explore modifications to the declarative memory architecture because the dataset is not large enough or rich enough to rigorously test modifications to the theory.
It also means that we do yet fully understand how well the ACT-R declarative memory theory scales when the number of chunks begins to approach the number of chunks that are actually stored in human declarative memory.

@@

Social media sites such as Twitter, Facebook, StackOverflow, Google+, and Yelp are composed almost entirely of human-created content.
The amount of user content on these sites is staggering, and continues to grow.
Users of Twitter's microblogging service, for example, generate half a billion tweets per day \parencite{TwitterReport2012}.
However, a single user of a social media site is most likely interested in only a small fragment of this large amount of information.
One general question to support the user in this domain is:
How can we quickly and effectively connect users to the content that they care about?

A growing number of users on social media sites are focusing on and looking for recently created information when performing a search \parencite{Jansen2011}.
In other words, users on social media sites want the fresh \emph{stream} of information about a particular topic.
This can certainly be seen in the main user interface views that these social media sites provide:
Twitter's homepage showing followees' tweets, Facebook's homepage showing friends' posts, and StackOverflow's daily digest of posts for particular tags.

@@

These two memory systems were compared across two domains where users generate tags for posts: Twitter and StackOverflow.
Twitter is a microblogging service where users create 140 character tweets and broadcast those messages out to the users who follow their posts.
A user is free to annotate important concepts or words in the tweet by preceding words with hashtags ``\#.''
After the tweet is created the hashtags become hyperlinks, and if clicked on will take a user to a summary feed of all of the tweets across Twitter that have used that hashtag.
So these user-created hashtags on Twitter help connect the tweet to other tweets about the same topic through common hashtag use.
Some recent and often used hashtags for Twitter are \emph{\#photography}, \emph{\#startup}, \emph{\#4change}, \emph{\#android}, and \emph{\#solar}.

@@

Some example tweets on Twitter are included in Figure \ref{figTweetExample}.
Note how users both add hashtags to the end of a tweet as a summary of the message and emphasize important concepts words in the tweet by turning them into hashtags.
That is, hashtags are both intermixed within the words of the tweet and added to the end of the message.

These memory retrieval models were also tested on tag use for StackOverflow posts.
StackOverflow is a question and answer site for computer programming where users ask programming-related questions and fellow members of the community provide answers.
A user with a programming-related question creates a post with their question and tags the post with a few specific programming-related tags. 
The tags chosen by a user on the StackOverflow site represent the primary topic of the question, such as a specific programming language, tool, software package, or framework.
Some example tags for StackOverflow are \emph{PHP}, \emph{Arrays}, \emph{MVC}, \emph{C\#}, and \emph{Common Lisp}.

@@

Also, studying user-based tag generation on these sites has relevant real-world application.
Models that can accurately predict the tags that users will generate on Twitter and StackOverflow can be used as the foundation for recommendations systems on these very popular sites.
These systems can also help newer users by recommending proper tags for their newest content.
On the StackOverflow site, for example, experts often subscribe to specific hashtags that interest them, and receive a daily digest of posts that are tagged with those hashtags.
Helping the user properly tag posts on StackOverflow ensures that right community sees the post, which greatly increases the chance that the question on the post will be answered quickly and correctly.
For Twitter, hashtags represent streams of information that are possibly interesting to the user.
Having a recommendation system that can suggest relevant hashtags to the user provides a way for the user to connect to and discover new information streams that interest them.

@@

The ACT-R declarative memory theory and the vector-based retrieval models will be described.
The Latent Semantic Analysis (LSA) theory is closely related to vector-based retrieval models, so an overview of LSA will be included.
The theory of the ACT-R and vector-based models will be compared.
Also, different tagging models have already been applied to various recommendation systems to tag user-generated content,
so these systems will be described and the models used will be related back to the Bayesian and vector-based declarative memory models.

@@

ACT-R \parencite{Anderson2007} is a cognitive architecture that formalizes how each cognitive process of the mind (e.g., memory, learning, visual and motor) interacts to produce behavior.
The declarative memory system is a component of that architecture that models the timing, learning, and forgetting processes that make up declarative memory storage and retrieval.
The equations that make up this system were derived from a rational analysis of long-term memory retrieval.
That is, given the task of retrieving a chunk of information from long-term declarative memory,
the current context (i.e., external and internal environment state), and past experience (i.e., prior memories and exposure), 
what is the optimal behavior (i.e., the optimal chunk to retrieve from memory)?
Using Bayesian reasoning, each chunk of information in declarative memory can be assigned a prior likelihood of needing retrieval again, given the prior history of exposure to the chunk.
These chunk prior probabilities are then adjusted for the current context, so that the posterior probabilities represent the likelihood that a chunk is needed, given prior odds and adjusted by current environment state.

@@

\begin{table}[!ht]
  \caption{ACT-R declarative memory model}
  \label{tabACTRModel}
  {\tabulinesep=1.2mm
    \begin{tabu}{ll}
      \hline
      Common Name &  Equation \\
      \hline
      Activation &	 	$A_{i} = B_{i} + \sum_{j \in c}^{} W_{j} S_{ji}$ \\
      Attentional Weight &	$W_{j} = \frac{W}{n}$ \\
      Base Level & 		$B_{i} = log \sum_{j=1}^{n} {t_{j}}^{-d}$ \\
      Constant Base Level &	$B_{i} = log \frac{p(i)}{p(\overline{i})}$ \\
      Strength of Association &	$S_{ji} = log \frac{p(i|j)}{p(i|\overline{j})} \approx log \frac{p(i|j)}{p(i)} = log \frac{NN_{ji}}{N_{Row}(j)N_{Col}(i)}$ \\
      Recall Probability &	$P_{i} = \left( 1 + e^{\frac{\tau - A_{i}}{s}} \right )^{-1}$ \\
      \hline
    \end{tabu}
  }
\end{table}

The total activation ($A_{i}$) for a chunk in declarative memory is a function of two components: base-level activation ($B_{i}$) and strength of association ($S_{ji}$).
The recall probability ($P_{i})$ that a chunk will be retrieved from memory increases with total activation ($A_{i}$).

@@

Pointwise Mutual Information (PMI) is another co-occurrence index that measures strength of association between two terms \parencite{Farahat2004}.
It is based on the co-occurrence count between pair-wise observations of terms, similar to ACT-R's strength of association.
The PMI index is presented as Equation \eqref{eqPMI}.

@@

@@

The majority of tag-recommendation systems use some measure of word pair associations to compute the likelihood of a tag, given current context.
As a applied concern, it is important that this measure scales well enough to handle the large volume of information present in online social media datasets.
Large social media datasets such as StackOverflow and Twitter contain on the order of millions to hundreds of millions of unique word pair associations between terms and tags.
If one wants to capitalize on all of the information present in the database to build the most accurate measure of relatedness between words and tags, then the association measure used must scale to this large size.

@@

The SNIF-ACT framework \parencites{Fu2007,Pirolli2003} is also based on the large-scale version of $S_{ji}$, and scaled well when working with a large-scale search query database for internet search.

As one of the largest tests, \textcite{Farahat2004} used the PMI index based on a 118GB local database of 10 million pages of the Stanford WebBase Project.
Even at this large dataset size, the ACT-R declarative memory theory still scaled well enough for reasonably fast retrieval times for $\mathit{PMI}_{ji}$ values.

@@

Latent Semantic Analysis \parencite{Landauer1997} (LSA) is an alternative representation of declarative memory.
Both ACT-R's Bayesian declarative memory model and LSA compute activations based on a co-occurrence matrix.
However, LSA performs further processing on that matrix in order to collapse words that have similar meaning.
The process starts with a set of N documents.
A word by document co-occurrence matrix is built by counting the number of times each word appears in each document.
The dimensionality of the full word frequency co-occurrence matrix is then reduced, while maintaining as much of the variance in the original full matrix as possible.
This process is similar to how factor analysis takes a data table and finds the most efficient way to represent that data with as few dimensions as possible.
The strength of association between two words is measured by the cosine of the two word vectors across the reduced dimension space.

@@

Although the LSA model has been shown to perform well as a way to measure similarity between two terms, one of the main criticisms is that it is a ``bag of words'' model.
This means that the model does not account for or pay attention to word order when building the base word by document matrix or consequently, when performing the SVD dimension reduction.
Other models that can account for word order have been shown to perform better than ``bag of words'' models on tasks that do not even appear to require grammatical structure such as word order to do well.
For example, \textcite{Jones2007} compared a vector-based model called BEAGLE directly to LSA, training and testing both models on the same corpus used in \textcite{Landauer1997}.
The context-only BEAGLE model (no word order) performed nearly equivalent to the LSA model at the task (\num{55.6}\% vs. \num{55.3}\%).
However, even in a multiple-choice domain where the task is to identify synonyms (very little grammatical structure in the task), adding word order when training the model improved task performance to \num{57.5}\%.
It seems reasonable to expect that task performance will improve much more on tasks where word order is used (e.g., generating a hashtag in the middle of a sentence when composing a tweet).

@@

Another issue with LSA is that the SVD matrix decomposition technique is computationally very expensive.
As the size of the original word by document matrix increases, the SVD computation becomes more and more time consuming.
If the domains where LSA is applied contained small enough datasets where SVD computation could be computed in a reasonable amount of time, then this would not be much of a concern.
However, with large human-created datasets such as Twitter, StackOverflow, and Wikipedia, the total size of the dataset far exceeds the maximum size that can even be computed by an LSA approach.
\textcite{Budiu2007} used the first six million pages of the Stanford WebBase corpus to compare performance between LSA and PMI on the TOEFL, and the LSA model could not be implemented with this size training dataset.
Instead, they compared the measures on the TOEFL by training on the smaller Touchstone Applied Science Associates (TASA) corpus created from \num{60527} samples of text from high school to early college textbooks.
This is one of the available corpora on the LSA website and is intended to summarize the general reading up to the 1st year of college \parencite{Budiu2007}.

When \textcite{Budiu2007} compared the simpler PMI (i.e., $S_{ji}$) measure to LSA on smaller TASA-sized datasets, LSA performed much better (23\% vs. 60\%).
This was one of the primary motivating reasons that LSA was developed in the first place, since reduced dimensionality model outperformed the full dimensionality model on TASA-sized datasets.
However, once the dataset surpasses in size what is computationally feasible with LSA, the much simpler PMI measure can still be computed, and the performance for PMI becomes on par and even surpasses that of LSA.
\textcite{Budiu2007} also trained the PMI measure on the much larger Stanford WebBase corpus and tested the measure on the TOEFL.
This PMI model increased accuracy from 23\% to 51\% when training the dataset on the smaller TASA compared to the larger Stanford dataset.
Further, when \textcite{Turney2001} trained the PMI measure on the even larger AltaVista index of 350 million webpages, the accuracy on the TOEFL increased to \num{73.75}\%, which is 10\% higher than the TASA-trained LSA.
This suggests that simpler models may be preferred to more complex (yet more accurate) models in large-scale domains, since the simpler models scale and can utilize all available information to generate predictions.

@@

The resonance and decoding processes are inversions of each other.
Usually both resonance and decoding will return very similar rank orderings of most likely chunks given context.
\textcite{Jones2007} recommends averaging the cosine values returned from each process for each chunk, and then rank ordering those averages to determine the chunk to retrieve.

@@

Both random permutations and LSA calculate word similarity on a reduced matrix.
For LSA, the original word by document matrix is compressed to a factor by document matrix and the number of factors are much less than the number of words.
For the StackOverflow and Twitter datasets, the number of documents are the number of different tags.

@@

This is a counterintuitive result.
It seems plausible that an SVD technique that takes into account word similarity when reducing the matrix should generate a reduced matrix that has more information than a randomly-compressed matrix.
This does not seem to be the case in testing however, as vector-based approaches like BEAGLE and random permutations perform just as well as the LSA when trained and tested on similar datasets.
Nonetheless, it has been suggested to perform an SVD reduction technique on a compressed random permutation vector \parencite{Kanerva2000}.
This approach might be worth investigating, as it would reduce computational requirements even further, and possibly improve performance by pooling co-occurrence counts from latent topics together.

@@

Vector-based models can be thought of as a compressed representation of the full co-occurrence matrix used for the ACT-R declarative memory system.
Each model computes an activation for each word to retrieve, given context and (for ACT-R's Bayesian model) prior knowledge.
Since both models produce activations for each chunk, one can be swapped out for another and used as the declarative memory component of the ACT-R infrastructure.
For example, \textcite{Rutledge2007} used a variant of the BEAGLE vector-based model as the DM component of ACT-R.
\textcite{Rutledge2008} showed that an ACT-R theory with a vector-based memory system can successfully model the fan effect, which is a psychological phenomena that ACT-R's current Bayesian DM system can model as well.

@@

Both vector-based and Bayesian models produce the fan effect since they both are based on a co-occurrence matrix.
However, vector-based models do further processing on that matrix, reduce its dimensionality, and use a correlation instead of a log odds value to compute the activation.
Nonetheless, vector-based models still produce the fan effect because at their core they are still computing activations based on co-occurrences.

@@

One strength of a vector-based model over ACT-R's Bayesian DM model is that word order can be naturally represented alongside unordered co-occurrence information.
With the ACT-R model, for example, it is unclear how to represent that a particular word appeared just before a hashtag. 
Although there are certainly ways to incorporate word order into the ACT-R DM model.
For example, an additional dimension could be added to the co-occurrence matrix that represents word position.
But adding word-order information to ACT-R's default Bayesian model is not common.

@@

The Netflix Prize \parencite{Bennett2007} is an example of a recommendation system where prior user behavior and current context was used to present the most relevant information to the user.
Netflix is an online subscription service where users can stream their favorite television episodes and movies.
Each user of the site is only interested in a small portion of the entire library of content that Netflix provides.
In order to provide a high quality user experience on the site, Netflix created a content recommendation service.
This tool recommends to the user movies and television episodes that most likely match their specific interests.

@@

Lucene \parencite{McCandless2010} is a commonly-used open source software tool that measures the similarity between a query and candidate documents in a database.
It uses the term frequency---inverse document frequency co-occurrence measure to quantify the similarity between a word in the query and a possible document to retrieve.
This measure is the product of two terms: the frequency count of the word in the document and the log inverse of the ratio of the number of documents that the word appeared in to the total number of documents.
The term frequency component measures the strength of association between a word and a document.
The inverse document frequency component measures how much information is contained within a specific word, independent of document.
ACT-R also uses two similar components to compute the activation of a word with a specific document.
The strength of association component of the ACT-R Bayesian co-occurrence model is a more sophisticated method for computing the term frequency component,
and ACT-R's attentional weight component of the declarative memory model can be used as a measure of the information content within each word.

@@

\textcite{Kuo2011} was the first to model tag use for StackOverflow posts.
His set of models were originally designed for next-word prediction for large text document collections.
\citeauthor{Kuo2011} examined the accuracy of three different commonly-used models for next-word prediction: a Bayesian model, a K-nearest-neighbors model, and a Latent Semantic Analysis (LSA) model.
In order to structure the models to work for StackOverflow tag prediction, the models would request the most-likely next word after a user created a post, and restrict the space of possible words to only tags.
He tested the set of next-word prediction models to generate predictions for the most likely tags that a user would choose when asking a question on the site.
Interestingly, the Bayesian co-occurrence model outperformed the more complicated and computationally expensive LSA (since it requires Singular Value Decomposition) and K-nearest-neighbors models.
This Bayesian model can accurately predict a user-chosen tag at 47\% accuracy.

@@

The hashtag recommendation system deployed for the Google+ microblogging service is probably the most ambitious model of user-created hashtags to date \parencite{GoogleKeynote2013}.
The user interface for this recommendation system for an example Google+ post is shown in Figure \ref{figGoogle+Hashtags}.

>>>>

\abstract{
  The growth of social media and user-created content on online sites provides unique opportunities to study models of declarative memory.
  The tasks of choosing a hashtag for a tweet and tagging a post on StackOverflow were framed as declarative memory retrieval problems.
  Two state-of-the-art cognitively-plausible declarative memory models were evaluated on how accurately they predict a user's chosen tags: an ACT-R based Bayesian model and a random permutation vector-based model.
  Millions of posts and tweets were collected, and both declarative memory models were used to predict Twitter hashtags and StackOverflow tags.
  The results show that past user behavior of tag use is a strong predictor of future behavior.
  Furthermore, past behavior was successfully incorporated into the random permutation model that previously used only context.
  Also, ACT-R's attentional weight term was linked to a common entropy-weighting natural language processing method used to attenuate low-predictor words.
  Word order was not found to be strong predictor of tag use, and the random permutation model performed comparably to the Bayesian model without including word order.
  This shows that the strength of the random permutation model is not in the ability to represent word order, but rather in the way in which context information is successfully compressed.
  Finally, model accuracy was moderate to high for the tasks, which supports the theory that choosing tags on StackOverflow and Twitter is primarily a declarative memory retrieval process. 
  The results of the large-scale exploration show how the architecture of the two memory models can be modified to significantly improve accuracy,
  and may suggest task-independent general modifications that can help improve model fit to human data in a much wider range of domains.
}

@@

The ACT-R cognitive architecture \parencite{Anderson2007} contains one of the leading theories of human cognitive processing.
The declarative memory component in ACT-R is a Bayesian statistical model.
This is the same type of co-occurrence-based Naive Bayes statistical model that is prominent in the more general machine learning field.
When these Bayesian models are used for machine learning, it is common to use a large number of observations to construct the co-occurrence matrix that is the foundation for the model.
The primary reason for this is that the stability of statistical models increase as sample size increases.

However, it is much less common in the cognitive modeling community to use a large number of observations to build up declarative memory when creating ACT-R models.
If a cognitive modeler is using the declarative memory component of ACT-R,
it is common to either construct the co-occurrences by hand, or use a small training set of data to initially build up the declarative memory system.
This approach can and has worked well.
Essentially, the modeler is only including the observations in declarative memory that are relevant to the specific task that they are studying.

@@

This has important implications for understanding declarative memory:
First, it is difficult to verify and explore modifications to the declarative memory architecture because the datasets are not large enough or rich enough to rigorously test modifications to the theory.
It also means that we do yet fully understand how well the ACT-R declarative memory theory scales when the number of chunks begins to approach the number of chunks that are actually stored in human declarative memory.

@@

Social media sites such as Twitter, Facebook, StackOverflow, Google+, and Yelp are composed almost entirely of human-created content.
The amount of user content on these sites is quite large, and continues to grow.
Users of Twitter's microblogging service, for example, generate half a billion tweets per day \parencite{TwitterReport2012}.
However, a single user of a social media site is most likely interested in only a small fragment of this large amount of information.
One general question to support the user in this domain is:
How can we quickly and effectively connect users to the content that they care about?

A growing number of users on social media sites are focusing on and looking for recently-created information when performing a search \parencite{Jansen2011}.
In other words, users on social media sites want the fresh \emph{stream} of information about a particular topic.
This can certainly be seen in the main user interface views that these social media sites provide:
Twitter's homepage showing followees' tweets, Facebook's homepage showing friends' posts, and StackOverflow's daily digest of posts for particular tags.

@@

These two memory systems were compared across two domains where users generate tags for posts: Twitter and StackOverflow.
Twitter is a microblogging service where users create 140 character ``tweets'' and broadcast those messages out to the users who follow their posts.
A user is free to annotate important concepts or words in the tweet by preceding words with ``\#,'' thereby making the word into a hashtag.
After the tweet is created the hashtags become hyperlinks, and if clicked on will take a user to a summary feed of all of the tweets across Twitter that have used that hashtag.
So these user-created hashtags on Twitter help connect the tweet to other tweets about the same topic through common hashtag use.
Some recent and often used hashtags for Twitter are \emph{\#photography}, \emph{\#startup}, \emph{\#4change}, \emph{\#android}, and \emph{\#solar}.

@@

Some example tweets on Twitter are included in Figure \ref{figTweetExample}.
Note how users both add hashtags to the end of a tweet as a summary of the message and emphasize important concepts words in the tweet by turning them into hashtags.
That is, hashtags can be both intermixed within the words of the tweet and added to the end of the message.

The two memory retrieval models were also tested on tag use for StackOverflow posts.
StackOverflow is a question and answer site for computer programming where users ask programming-related questions and fellow members of the community provide answers.
A user with a programming-related question creates a post with their question and tags the post with a few specific programming-related tags. 
The tags chosen by a user on the StackOverflow site represent the primary topic of the question, such as a specific programming language, tool, software package, or framework.
Some example tags for StackOverflow are \emph{PHP}, \emph{Arrays}, \emph{MVC}, \emph{C\#}, and \emph{Common Lisp}.

@@

Also, studying user-based tag generation on these sites has relevant real-world application.
Models that can accurately predict the tags that users will generate on Twitter and StackOverflow can be used as the foundation for recommendations systems on these popular sites.
These systems can also help newer users by recommending proper tags for their newest content.
On the StackOverflow site, for example, experts often subscribe to specific hashtags that interest them, and receive a daily digest of posts that are tagged with those hashtags.
Helping the user properly tag posts on StackOverflow ensures that right community sees the post, which greatly increases the chance that the question on the post will be answered quickly and correctly.
For Twitter, hashtags represent streams of information that are possibly interesting to the user.
Having a recommendation system that can suggest relevant hashtags to the user provides a way for the user to connect to and discover new information streams that interest them.
Also, by suggesting relevant hashtags the system can help make sure a user's content is seen by other users who are interested in the same type of information.

@@

The ACT-R declarative memory theory and the vector-based retrieval models will be described.
The Latent Semantic Analysis (LSA) theory is closely related to vector-based retrieval models, so an overview of LSA will also be included.
The theories underlying the ACT-R and vector-based models will be compared.
Also, different tagging models have already been applied to various recommendation systems to tag user-generated content.
Many of these models are based on the same Bayesian statistical framework that is the foundation of the ACT-R declarative memory theory.
So these systems will be described and the models used will be discussed within the context of Bayesian statistical models, LSA approaches, and the ACT-R declarative memory theory.

@@

ACT-R \parencite{Anderson2007} is a cognitive architecture that formalizes how each cognitive process of the mind (e.g., memory, learning, visual and motor) interacts to produce behavior.
The declarative memory system is a component of that architecture that models the timing, learning, and forgetting processes that make up declarative memory storage and retrieval.
The equations that make up this system were derived from a rational analysis of declarative memory retrieval.
That is, given the task of retrieving a chunk of information from declarative memory,
the current context (i.e., external and internal environment state), and past experience (i.e., prior memories and exposure), 
what is the optimal behavior (i.e., the optimal chunk to retrieve from memory)?
Using Bayesian reasoning, each chunk of information in declarative memory can be assigned a prior likelihood of needing retrieval again, given the prior history of exposure to the chunk.
These chunk prior probabilities are then adjusted for the current context, so that the posterior probabilities represent the likelihood that a chunk is needed, given prior odds and adjusted by current environment state.

@@

\begin{table}[!ht]
  \caption{ACT-R declarative memory model}
  \label{tabACTRModel}
  {\tabulinesep=1.2mm
    \begin{tabu}{ll}
      \hline
      Common Name &  Equation \\
      \hline
      Activation &	 	$A_{i} = B_{i} + \sum_{j \in c}^{} W_{j} S_{ji}$ \\
      Attentional Weight &	$W_{j} = \frac{W}{n}$ \\
      Base Level & 		$B_{i} = log \sum_{j=1}^{n} {t_{j}}^{-d}$ \\
      Constant Base Level &	$B_{i} = \frac{p_{i}}{1-p_{i}}$ \\
      Strength of Association &	$S_{ji} = log \frac{p(i|j)}{p(i|\overline{j})} \approx log \frac{p(i|j)}{p(i)} = log \frac{NN_{ji}}{N_{Row}(j)N_{Col}(i)}$ \\
      Recall Probability &	$P_{i} = \left( 1 + e^{\frac{\tau - A_{i}}{s}} \right )^{-1}$ \\
      \hline
    \end{tabu}
  }
\end{table}

The total activation ($A_{i}$) for a chunk in declarative memory is a function of two components: base-level activation ($B_{i}$) and strength of association ($S_{ji}$).
The recall probability ($P_{i})$ that a chunk will be retrieved from memory increases with total activation ($A_{i}$).
Each term in Table \ref{tabACTRModel} will be discussed in greater detail to follow.

@@

Pointwise Mutual Information (PMI) is another co-occurrence index that measures strength of association between two terms \parencite{Farahat2004}.
It is based on the co-occurrence count between pair-wise observations of terms, similar to ACT-R's strength of association.
The index has been used to accurately and efficiently measure strength of association information for large-scale datasets \parencite{Budiu2007,Farahat2004}.
The PMI index is presented as Equation \eqref{eqPMI}.

@@

The attentional weight term allows for some contextual components to be weighted more than others when the total $S_{ji}$ activation is computed.
This reflects the fact that the declarative memory system has limited attentional resources ($W$) that must be spread across all of the items in context.
Most ACT-R models simply weight each contextual element equally (e.g., all words in a StackOverflow post have equal weight), so each word's attentional weight is simply normalized by the number of items in context.

@@

The majority of tag-recommendation systems use some measure of word pair associations to compute the likelihood of a tag, given current context.
As an applied concern, it is important that this measure scales well enough to handle the large volume of information present in online social media datasets.
Large social media datasets such as StackOverflow and Twitter contain on the order of millions to hundreds of millions of unique word pair associations between terms and tags.
If one wants to capitalize on all of the information present in the database to build the most accurate measure of relatedness between words and tags, then the association measure used must scale to this large size.

@@

The SNIF-ACT framework \parencites{Fu2007,Pirolli2003} was developed to predict user navigation of web-based hyperlinks (e.g., choosing between a set of results from a search query).
The theory uses the large-scale version of $S_{ji}$ to measure the relatedness between a user's current goals and the current context.
It was shown to scale well when working with a large-scale search query database for internet search.

As one of the largest tests for the PMI index, \textcite{Farahat2004} generated the co-occurrence matrix from a 118GB local database of 10 million pages of the Stanford WebBase Project.
\citeauthor{Farahat2004} compared strength of association values using the PMI index and similarity ratings derived from human raters.
Even at this large dataset size, the ACT-R declarative memory theory still scaled well enough for reasonably fast retrieval times for $\mathit{PMI}_{ji}$ values.

@@

Latent Semantic Analysis \parencite{Landauer1997} (LSA) is an alternative representation of declarative memory.
Both ACT-R's Bayesian declarative memory model and LSA compute activations based on a co-occurrence matrix.
However, LSA performs further processing on that matrix in order to collapse words that have similar meaning.
The process starts with a set of N documents.
A word by document co-occurrence matrix is built by counting the number of times each word appears in each document.
The dimensionality of the full word frequency co-occurrence matrix is then reduced, while maintaining as much of the variance in the original full matrix as possible.
This process, described in more detail to follow, is similar to how factor analysis takes a data table and finds the most efficient way to represent that data with as few dimensions as possible.
The strength of association between two words is measured by the cosine of the two word vectors across the reduced dimension space.

@@

Although the LSA model has been shown to perform well as a way to measure similarity between two terms, one of the main criticisms is that it is a ``bag of words'' model.
This means that the model does not account for or pay attention to word order when building the base word by document matrix or consequently, when performing the SVD dimension reduction.
Other models that can account for word order have been shown to perform better than ``bag of words'' models on tasks that do not even appear to require grammatical structure such as word order to do well.
For example, \textcite{Jones2007} compared a vector-based model called BEAGLE directly to LSA, training and testing both models on the same corpus used in \textcite{Landauer1997}.
The context-only BEAGLE model (no word order) performed nearly equivalent to the LSA model at the TOEFL task (\num{55.6}\% vs. \num{55.3}\%).
However, even in a multiple-choice domain where the task is to identify synonyms (very little grammatical structure in the task), adding word order when training the model improved task performance to \num{57.5}\%.
It seems reasonable to expect that task performance will improve much more on tasks where word order is used (e.g., generating a hashtag in the middle of a sentence when composing a tweet).

@@

Another issue with LSA is that the SVD matrix decomposition technique is computationally expensive.
As the size of the original word by document matrix increases, the SVD computation becomes more and more time consuming.
If the domains where LSA is applied contained small enough datasets where SVD computation could be computed in a reasonable amount of time, then this would not be much of a concern.
However, with large human-created datasets such as Twitter, StackOverflow, and Wikipedia, the total size of the dataset far exceeds the maximum size that can even be computed by an LSA approach.
\textcite{Budiu2007} used the first six million pages of the Stanford WebBase corpus to compare performance between LSA and PMI on the TOEFL, and the LSA model could not be implemented with this size training dataset.
Instead, they compared the measures on the TOEFL by training on the smaller Touchstone Applied Science Associates (TASA) corpus created from \num{60527} samples of text from high school to early college textbooks.
This is one of the available corpora on the LSA website and is intended to summarize the general reading up to the 1st year of college \parencite{Budiu2007}.

When \textcite{Budiu2007} compared the simpler PMI (i.e., $S_{ji}$) measure to LSA on smaller TASA-sized datasets, LSA performed much better (23\% vs. 60\%).
This was one of the primary motivating reasons that LSA was developed in the first place, since reduced dimensionality model outperformed the full dimensionality model on TASA-sized datasets.
However, once the dataset surpasses in size what is computationally feasible with LSA, the much simpler PMI measure can still be computed, and the performance for PMI becomes on par and even surpasses that of LSA.
\textcite{Budiu2007} also trained the PMI measure on the much larger Stanford WebBase corpus and tested the measure on the TOEFL.
This PMI model increased accuracy from 23\% to 51\% when training the dataset on the smaller TASA compared to the larger Stanford dataset.
Further, when \textcite{Turney2001} trained the PMI measure on the even larger AltaVista index of 350 million webpages, the accuracy on the TOEFL increased to \num{73.8}\%, which is 10\% higher than the TASA-trained LSA.
This suggests that simpler models may be preferred to more complex (yet more accurate) models in large-scale domains, since the simpler models scale and can utilize all available information to generate predictions.

@@

The resonance and decoding processes are inversions of each other.
Usually both resonance and decoding will return similar rank orderings of most likely chunks given context.
\textcite{Jones2007} recommends averaging the cosine values returned from each process for each chunk, and then rank ordering those averages to determine the chunk to retrieve.

@@

Both random permutations and LSA calculate word similarity on a reduced matrix.
For LSA, the original word by document matrix is compressed to a factor by document matrix and the number of factors is much less than the number of words.
For the StackOverflow and Twitter datasets, the number of documents are the number of different tags.

@@

This is a counterintuitive result.
It seems plausible that an SVD technique that takes into account word similarity when reducing the matrix should generate a reduced matrix that has more information than a randomly-compressed matrix.
This does not seem to be the case in testing however \parencites{Sahlgren2008,Jones2007}.
Nonetheless, \textcite{Kanerva2000} suggested to perform an SVD reduction technique on a compressed random permutation vector.
This approach might be worth investigating, as it would reduce computational requirements even further, and possibly improve performance by pooling co-occurrence counts from latent topics together.

@@

Vector-based models can be thought of as a compressed representation of the full co-occurrence matrix used for the ACT-R declarative memory system.
Each model computes an activation for each word to retrieve, given context and (for ACT-R's Bayesian model) prior knowledge.
Since both models produce activations for each chunk, one can be swapped out for another and used as the declarative memory component of the ACT-R infrastructure.
For example, \textcite{Rutledge2007} used a variant of the BEAGLE vector-based model as the DM component of ACT-R.
\citeAuthor{Rutledge2008} showed that an ACT-R theory with a vector-based memory system can successfully model the fan effect.

@@

Both vector-based and Bayesian models produce the fan effect since they both are based on co-occurrences.
However, vector-based models do further processing on that matrix, reduce its dimensionality, and use a correlation instead of a log odds value to compute the activation.
Nonetheless, vector-based models still produce the fan effect because at their core they are still computing activations based on co-occurrences.

@@

One strength of a vector-based approach over ACT-R's Bayesian DM model is that word order can be naturally represented alongside unordered co-occurrence information.
With the ACT-R model, for example, it is unclear how to represent that a particular word appeared just before a hashtag. 
To represent word order in a Bayesian model, an additional dimension could be added to the co-occurrence matrix that represents word position.
However, adding word-order information to ACT-R's default Bayesian declarative memory model is not common.
Instead, when word order information is used to model reading for example, production rules are defined so that words are read from left to right (e.g., \textcite{Lewis2006}).

@@

The Netflix Prize \parencite{Bennett2007} is an example of a recommendation system where prior user behavior and current context was used to present the most relevant information to the user.
Netflix is an online subscription service where users can stream television episodes and movies.
Each user of the site is only interested in a small portion of the entire library of content that Netflix provides.
In order to provide a high quality user experience on the site, Netflix created a content recommendation service.
This tool recommends to the user movies and television episodes that most likely match their specific interests.

@@

Lucene \parencite{McCandless2010} is a commonly-used open source software tool that measures the similarity between a query and candidate documents in a database.
It uses the term frequency-inverse document frequency co-occurrence measure to quantify the similarity between a word in the query and a possible document to retrieve.
This measure is the product of two terms: the frequency count of the word in the document and the log inverse of the ratio of the number of documents that the word appeared in to the total number of documents.
The term frequency component measures the strength of association between a word and a document.
The inverse document frequency component measures how much information is contained within a specific word, independent of document.

@@

\textcite{Kuo2011} was the first to model tag use for StackOverflow posts.
His set of models were originally designed for next-word prediction for large text document collections.
\citeauthor{Kuo2011} examined the accuracy of three different commonly-used models for next-word prediction: a Bayesian model, a K-nearest-neighbors model, and a Latent Semantic Analysis (LSA) model.
In order to structure the models to work for StackOverflow tag prediction, the models would request the most-likely next word after a user created a post, and restrict the space of possible words to only tags.
He tested the set of next-word prediction models to generate predictions for the most likely tags that a user would choose when asking a question on the site.
Interestingly, the Bayesian co-occurrence model outperformed the more complicated and computationally expensive LSA and K-nearest-neighbors models.
This Bayesian model can accurately predict a user-chosen tag at 47\% accuracy.

@@

The hashtag recommendation system deployed for the Google+ social media site is probably the most ambitious model of user-created hashtags to date \parencite{GoogleKeynote2013}.
The user interface for this recommendation system for an example Google+ post is shown in Figure \ref{figGoogle+Hashtags}.
